---
title: "Data transformation"
author: "Yifan Jin"
date: "09/02/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Visualisation is an important tool for insight generation, but it is rare that you get the data in exactly the right form you need. We may need to create, summarises, rename, reorder vairables.

### Prerequisites

```{r,warning=FALSE}
# install.packages("nycflights13")
library(nycflights13)
library(tidyverse)
```

### nycflights13

```{r}
flights
```

`int`: intergers

`dbl`: doubles, or real numbers

`chr`: character vectors, or strings

`dttm`: date-times (a date+ a time)

`lgl`: logical, vectors that contain only `TRUE` or `FALSE`

`fctr`: factors, which R uses to represent categorical variables with fixed possible values

`date`: dates

### dplyr basics

`filter()`: pick observations by their values

`arrange()`: reorder the rows

`select()`: pick variables by their names

`mutate()`: create new variables with functions of existing variable

`summarise()`: collapse many values down to single summary

All verbs work similarly:

1. The first argument is a data frame.

2. The subsequent arguments describe what to do with the data frame, using the variable names (without quotes)

3. The result is a new data frame


## Filter rows with filter()

`filter()` allows you to subset observations based on their values. The first argument is the name of the data frame. The second and subsequent arguments are the expressions that filter the data frame.

```{r}
filter(flights,month==1,day==1)
```

```{r}
jan1<-filter(flights,month==1,day==1)

(dec25<-filter(flights,month==12,day==25))
# brackets that we can print the result
```

### Comparisons

Comparison operator: `>`, `>=`, `<`,`<=`, `!=` and `==`

```{r}
# filter(flights,month=1) This is false since we need to use ==
sqrt(2)^2==2
1/49*49==1
# This result is surprising, when we compare two numbers, one is floating, one is integer

#Instead of relying on `==`, we use near()

near(sqrt(2)^2,2)
near(1/49*49,1)
```

### Logical operators
`&`: and

`|`: or

`!`: not

```{r}
filter(flights,month==11|month==12)

# or

nov_dec<-filter(flights,month %in% c(11,12))
```

### Missing values
```{r}
NA>5

10==NA

NA+10

NA/2
```

```{r}
NA==NA

X<-NA
Y<-NA
X==Y
```

```{r}
df<-tibble(x=c(1,NA,3))
filter(df,x>1)
filter(df,is.na(x)|x>1)
```

### Exercises

**1.Find all flights that had an arrival delay of two or more hours**

**Answer**

```{r}
filter(flights,arr_delay>120)
```

**2. Find all flights that Flew to Houston (`IAH` or `HOU`)**

**Answer**

```{r}
filter(flights,dest=="IAH"|dest=="HOU")
```

**3. Find all flights that were operated by United, American, or Delta**

**Answer**

```{r}
filter(flights,carrier=="UA"|carrier=="AA"|carrier=="DL")
```

**4. Departed in summer (July, August, and September)**

**Answer**

```{r}
filter(flights,month==7|month==8|month==9)
```

**5. Arrived more than two hours late, but didn't leave late**

```{r}
filter(flights,arr_delay>120|dep_delay<=0)
```

**6.Were delayed by at least an hours late, but made up over 30 minutes in flight**

**Answer**

```{r}
filter(flights,arr_delay<dep_delay-30|dep_delay>=60)
```

**7. Departed between midnight and 6am (inclusive)**

**Answer**
```{r}
filter(flights,dep_time>=000&dep_time<=600)
```

**8. Another useful dplyr filtering helper is `between()`. What does it do? **

**Answer**

```{r}
filter(flights,between(dep_time,000,600))
```

```{r}
filter(flights,is.na(dep_time))
```

## Arrange rows with arrange()

`arrange()` works similarly to `filter()` except that instead of selecting rows, it cahnges their order.

```{r}
arrange(flights,year,month,day)
```

Use `desc()` to re-order by a column in descending order:

```{r}
arrange(flights,desc(dep_delay))
```

Missing values are always sorted at the end:

```{r}
df<-tibble(x=c(5,2,NA))
arrange(df,x)

arrange(df,desc(x))
```

### Exercises

**1. How could you use `arrange()` to sort all missing values to the start? (hint: use `is.na()`)**

**Answer**

```{r}
arrange(df,desc(is.na(x)))
```

**2. Sort `flights` to find the most delayed flights. Find the flights that left earliest**

**Answer**

```{r}
arrange(flights,desc(dep_delay))
arrange(flights,dep_delay)
```

**3. Sort `flights` to find the fastest (highest speed) flights.**

**Answer**

```{r}
arrange(flights,desc(distance/air_time))
```

**4. Which flights travelled the farthest? Which travelled the shortest?**

**Answer**

```{r}
arrange(flights,desc(distance))
arrange(flights,distance)
```

## Select columns with select()

It is not uncommon to get datasets with hundreds or even thousands of variables. In this case, the first challenge is to narrowing in on the variables you are actually interested in. We use `select()`.

```{r}
select(flights,year,month,day) #directly type columns that want to stay
select(flights,year:day) # use : to select variables between left and right
select(flights,-(year:day)) # delete the columns that dont want
```


There are a number of helper functions you can see within `select()`:

1. `starts_with("abc")`: matches names that begin with "abc"

2. `ends_with("xyz")`: matches names that end with "xyz"

3. `contains("ijk")`:macthes names that contain "ijk"

4. `matches("(.)\\1")`: selects variables that match a regular expression. This one macthes any variables that contain repeated characters. You will learn more about regular expressions in strings.

5. `num_range("x",1:3)`: matches `x1`,`x2` and `x3`.

`select()` can be used to rename variables, but it is rarely useful because it drops all variables not explicitly mentioned. Instead, use`rename()`, which is a variant of `select()` that keeps all the variables that are not explicitly mentioned:

```{r}
rename(flights,tail_num=tailnum)
```

Another option is to use `select()` in conjunction with the `everything()` helper. 

```{r}
select(flights,time_hour,air_time,everything())
```

### Exercises

**1. Brainstorm as many ways as possible to select `dep_time`, `dep_delay`,`arr_time`, and `arr_delay` from `flights`**

**Answer**

```{r}
# 5!=5*4*3*2*1
```

**2. What happens if you include the name of a variable multiple times in a `select()` call?**

**Answer**

```{r}
select(flights,distance,distance)
```

Only one column will be showed.

**3. What does the `one_of()` function do? Why might it be helpful in conjunction with this vector?**

**Answer**

```{r}
vars<-c("year","month","day","dep_delay","arr_delay")
select(flights,one_of(vars))
# one_of(): Matches variable names in a character vector.
```

**4. Does the result of running the following code surprise you? How do the select helpers deal with case by default? How can you change that default?**

**Answer**

```{r}
select(flights,contains("TIME"))
```

Including all variables whcih name contains time.

## Add new variables with mutate()

`mutate()`: building new variable with existing variables

```{r}
flights_sml<-select(flights,year:day,
                    ends_with("delay"),
                    distance,
                    air_time)

mutate(flights_sml,
       gain=dep_delay-arr_delay,
       speed=distance/air_time*60)
```


```{r}
# You can refer to columns that you have just created
mutate(flights_sml,
       gain=dep_delay-arr_delay,
       hours=air_time/60,
       gain_per_hours=gain/hours)
```


If you only want to keep the new variables, use `transmute()`

```{r}
transmute(flights,
          gain=dep_delay-arr_delay,
          hours=air_time/60,
          gain_per_hour=gain/hours)
```

### Useful creation functions

There are many functions for creating new variables that you can use with `mutate()`. The key property is that the function must be vectorised: it must take a vector of values as input, return a vector with the same number of values as output. There is no way to list every possible function that you might use, but there is a selection of functions that are frequently useful:

1. Arithmetic operators: `+`,`-`,`*`,`/`,`^`. These are all vectorised using the so called "recycling rules". `x/sum(x)`- calculate the proportion of a total, `y-mean(y)`- calculate the difference from the mean

2. Modular airthmetic: `%/%` (integer division) and `%%` (remainder), where `x==y*(x%/%y)+(x%%y)`. Modular arithmetic is a handy tool because it allows you to break integers up into pecies.

```{r}
transmute(flights,
          dep_time,
          hour=dep_time%/% 100,
          minute=dep_time %%100)
```

3. Cumulative and rolling aggregates: R provides functions for running sums, products, mins and maxes: `cumsum()`, `cumprod()`, `cummin()`, `cummax()`; and dplyr provides `cummean()` for cumulative means.

```{r}
x<-c(1,2,3,4,5,6,7,8,9,10)
cumsum(x)
cummean(x)
```

4. Logical comparisons, `<`,`<=`,`>`,`>=`,`!=`, and `==`, which you learned earlier.

5. Ranking: there are a number of ranking functions, but you should start with `min_rank()`, use desc to get largest to smallest.

```{r}
y<-c(1,2,2,NA,3,4)
min_rank(y)
min_rank(desc(y))
```

## Grouped summarises with summarise()


