---
title: "Basic workflow"
author: "Yifan Jin"
date: "09/02/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Coding basics

```{r}
1/200*30
sin(pi/2)
# You can also assign a value to a variable
x<- 3*4
# object_name<-value
```

## What's in a name?
We need to use `_` and `.` 
```{r}
# i_use_snake_case
# otherPeopleUseCameCase
# some.people.use.periods
# And_aFew.People_RENOUNCEconvention
```

```{r}
# You can inspect an object by typing its name
x
# make another assignment
this_is_a_really_long_time<-2.5
this_is_a_really_long_time<-3.5
```

```{r}
# Make another assignment
r_rocks<-2^3
# r_rock and R_rocks is not gonna be work
```

## Calling functions
```{r}
# function_name(arg1=var1,arg2=val2)
```

Remeber when you deisgn a function your environment will help you store the function and also the variable you assigned

### Exercise
**1. Why does this code not work?**
```{r}
my_variable<-10
# my_var1able
```
**Answer**

You should change my_var1able to my variable...

**2.Tweak each of the following R commands so that they run correctly**

**Answer**
```{r}
library(tidyverse)
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ,y=hwy))
filter(mpg,cyl==8)
filter(diamonds,carat>3)
# completed
```

**3. Press Alt + Shift + K. What happens? How can you get to the same place using the menus?**

**Answer**

We can see all the command we can use.


