---
title: "Exploratory Data Analysis"
author: "Yifan Jin"
date: "15/02/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This chapter will show you heo to use visualisation and transformation to explore your data in a systematic way, a task that statisticians call exploratory data analysis, or EDA for short. EDA is an iterative cycle.

1. Generate questions about your data

2. Search for answers by visualising, transforming, and modelling your data.

3. Use what you learn to refine your questions and/or generate new questions

EDA is not a formal process with a strict set of rules. During the initial phases of EDA you should feel free to investigate every idea that occurs to you. Some of them will pan out, and some will be dead ends.

To do data cleaning, you will need to deploy all the tools of EDA: visualisation, transformation, and moelling.

### Prerequisites

In this chapter, we will combine what you have learned about dplyr and ggplot2 to interactively ask questions, answer them with data, and then ask new questions.

```{r}
library(tidyverse)
```

## Questions

Your goal during EDA is to develop an understanding of your data. The easiest way to do this is to use questions as tools to guide your investigation. When you ask a question, the question focuses your attention on a specific part of your dataset and helps you decide which graphs, models, or transformations to make.

EDA is fundamentally a creative process. And like most creative processes, the key to asking quality questions is to generate of questions. It is difficult to ask revealing questions at the start of your analysis because you do not know what insights are contained in your dataset. On the other hand, each new question taht you ask will expose you to a new aspect of data and increase your chance of making a discovery. 

There is no rule about which questions you should ask to guide your research. 
However, two types of questions will always be useful for making discoveries within your data.

**1. What type of variation occurs within my variables?**

**2. What type of covariation occurs between my variables?**

The rest of this part will look at these two questions.

1. Explain what variation and covariation are

2. several ways to answer each question.

`Variable`: A variable is a quantity, quality, or property that you can measure

`Value`: A value is the state of a variable when you can measure it. The value of a variable may change from measurement to measurement.

`Observation`: A observation is a set of measurements made under similar conditions (you usually make all of the measurements in an observation at the same time and on the same object). An observation will contain several values, each associated will a different variable. An observation is a data point.

`Tabular data`: Tabular data is a set of values, each associated with a variable and an observation. Tabular data is tidy if each value is placed in its own cell, each variable in its own column, and each observation in its own row.

## Variation

Variation is the tendency of the values of a variable to change from measuremnet to measurement. You can see variation easily in real life. Each variable has its own pattern of variation, which can reveal interesting information. The best way to understand that pattern is to visualse the distribution of the variable values.

### Visualising distributions

A variable is `categorical` if it can only take one of a small set of values. In R, categorical variables are usually saved as factors or character vectors. To examine the distribution of a categorical variable, use a bar cahrt

```{r}
ggplot(data=diamonds)+
  geom_bar(mapping=aes(x=cut))
```

The height of the bars displays how many observations occurred with each x value. You can compute these values manually with: `dplyr::count()`

```{r}
diamonds %>%
  count(cut)
```

A variable is `continuous` if it can take any of an infinite set of ordered values. Numbers and date-times are two examples of continuous variables. To examine the distribution of a continuous variable, we use histogram

```{r}
ggplot(data=diamonds)+
  geom_histogram(mapping=aes(x=carat),binwidth = 0.5)
```

You can compute this by hand by combining `dplyr::count()` and `ggplot2::cut_width()`:

```{r}
diamonds %>%
  count(cut_width(carat,0.5))
```

